---
title: "Hayman Fire Recovery"
author: "Megan Sears"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)
library(ggrepel)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

knitr::opts_knit$set(root.dir='..')
```


```{r dataread, warning=F,message=F}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files('../data',full.names=T)


#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')


ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))


```




## Question 1) 

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlation as a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months

```{r}
#make wide data format
full_wide <- spread(data=full_long,key='data',value='value') %>%
  filter_if(is.numeric,all_vars(!is.na(.))) %>% 
  mutate(month = month(DateTime),
         year = year(DateTime))
  

full_summer <- full_wide %>%
  filter_if(is.numeric,all_vars(!is.na(.)))
  

ggplot(full_summer, aes(ndmi, ndvi, color=site)) +
  geom_point()

```


## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?
 
```{r}
#ndsi for burned and unburned combined
ndsi_win <- full_wide %>%
  filter(month %in% c(1,2,3,4)) %>%
  group_by(year) %>%
  summarize(si_avg = mean(ndsi))

#ndvi for burned and unburned combined
ndvi_summ <- full_wide %>% 
  filter(month %in% c(6,7,8)) %>%
  group_by(year) %>%
  summarize(vi_avg = mean(ndvi))

#join the 2 above
q2_pt1 <- inner_join(ndsi_win, ndvi_summ, by=c("year"))

#plot showing comparison of ndsi and ndvi for burn and unburn combined
ggplot(q2_pt1, aes(x=si_avg, y=vi_avg)) +
  geom_point(size=4) 

#ndsi grouped by year and site
ndsi_q2 <- full_wide %>%
    filter(month %in% c(1,2,3,4)) %>%
  group_by(year, site) %>%
  summarize(si_avg = mean(ndsi))

#ndvi grouped by year and site
ndvi_q2 <- full_wide %>%
  filter(month %in% c(6,7,8)) %>%
  group_by(year, site) %>%
  summarize(vi_avg = mean(ndvi))

#ndvi and ndsi comparison by year and site for pre and post burn
#this answers q3
q2 <- inner_join(ndsi_q2, ndvi_q2, by=c("year", "site")) %>%
  mutate(fire = if_else(year > 2003, "yes", "no"))

ggplot(q2, aes(x=si_avg, y=vi_avg, color=site, shape=fire)) +
  geom_point(size=4)

#adding facet wrap to break out the burn and unburned - this answers q3
ggplot(q2, aes(x=si_avg, y=vi_avg, shape=fire)) +
  geom_point() +
  facet_wrap(~site)


```
 
## Q3

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 

Answers included in the q2 chunk.

## Question 4

What month is the greenest month on average? 

```{r}
#greenest month
mean_summary <- full_wide %>%
  group_by(month) %>% 
  summarize(moavg_ndvi = mean(ndvi),
            moavg_ndsi = mean(ndsi),
            moavg_ndmi = mean(ndmi))

mean_summary

```



## Question 5) 

What month is the snowiest on average?


## Bonus Question: Redo all problems with `spread` and `gather` using modern tidyverse syntax. 

```{r}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files('../data',full.names=T)


#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')

ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>% 
  pivot_longer(!c(DateTime, data), names_to="condition", values_to="value" ) %>%
  filter(!is.na(value))

#make wide data format
full_wide <- full_long %>%
  pivot_wider(names_from = data, values_from = value) %>% 
  filter_if(is.numeric,all_vars(!is.na(.))) %>% 
  mutate(month = month(DateTime),
         year = year(DateTime))

```



## Bonus Question: Use Climage Engine to pull the same data for the assignment, but updated with 2020/2021 data.




